### 2018-07-18 Introduction/Introduction.tex
First commit

\documentclass[letterpaper]{article}
\usepackage{graphicx}
\usepackage{subfigure}
\graphicspath{ {images/} }
\usepackage{enumerate}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{minted}
\usepackage{tcolorbox}
\usepackage[margin=1in]{geometry}
\usepackage{float}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{indentfirst}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
\renewcommand{\arraystretch}{1.2}

\begin{document}

\title{Master's Thesis}
\author{Sun Myung Park}
\date{}

\maketitle

\section*{Introduction}

\end{document}