### 2018-08-08 Introduction/Introduction.tex
Introduction work

\documentclass[letterpaper]{article}
\usepackage{graphicx}
\usepackage{subfigure}
\graphicspath{ {images/} }
\usepackage{enumerate}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{minted}
\usepackage{tcolorbox}
\usepackage[margin=1in]{geometry}
\usepackage{float}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{indentfirst}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
\renewcommand{\arraystretch}{1.2}

\begin{document}

\title{Master's Thesis}
\author{Sun Myung Park}
\date{}

\maketitle

\section*{Introduction}

With the world population slated to continue growing for the rest of the 21st century, our energy consumption is likely to follow suit. The bulk of energy production today comes from burning fossil fuels. However, the resulting release of carbon dioxide, a greenhouse gas, has contributed to rising temperatures and more severe weather phenomena around the globe. It is imperative that we eliminate the major source of greenhouse gases in fossil fuel burning power plants as quickly as possible to prevent worsening climate change.

\end{document}