### 2020-07-03 introduction.tex
editing

\section{Background and Motivation}

Greenhouse gas emission from human activities is the main cause of climate
change, a global phenomenon characterized by shifting weather patterns and
rising temperatures. The concentration of CO$_2$, a greenhouse gas, in the
atmosphere has been rising steadily throughout modern human history
\cite{us_department_of_commerce_global_nodate}. Climate change has dire
consequences on human health and safety due to extreme weather events and the
overall impact on food production \cite{noauthor_global_nodate}. As a result,
urgent measures are necessary to limit greenhouse gas emissions, namely CO$_2$
emissions from electricity generation.

Electricity generation from burning fossil fuels represents the
largest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019});
replacing it with low-carbon
alternatives would curb a large fraction of emissions. Nuclear power is
a viable low-carbon replacement for burning fossil fuels.
Other low-carbon energy sources include solar-, wind-, and hydro-power. Each
energy source has its own set of advantages and disadvantages; nuclear power's
advantage over renewable energy sources is that it provides consistent
base-load power independent of weather and geographical location
\cite{petti_future_2018}. Employing a diverse mix of these energy sources
ensures energy security and reliability in our transition
towards a low-carbon future \cite{petti_future_2018}.

The world would have to ramp up the current rate of reactor deployments to
displace some of the
presently large share of energy production from fossil fuel power plants.
However, several obstacles stand in the way of mass reactor deployments. These
include perceived safety risks, sustainability concerns, nuclear proliferation
risks, and the ability to compete economically with other sources of energy
\cite{massachusetts_institute_of_technology_future_2003}. A potential solution
to the aforementioned issues is the \gls{MSR} concept, one of six advanced
reactor designs selected by the Generation IV International Forum
\cite{gif_technology_2002} for continued research and development.

The primary coolants in MSRs consist of a molten salt mixture
with fissile and/or fertile material directly dissolved in the coolant.
MSRs possess an inherently robust safety feature in the strongly negative fuel
temperature coefficient of reactivity. Some designs can also incorporate the
thorium fuel cycle for improved sustainability arising from the use of
abundant natural thorium resources and reduced transuranic waste. The
latter also reduces economical costs
associated with long-term nuclear waste storage. In addition, the ability to
operate at near atmospheric pressures eliminates the need for a thick pressure
vessel and drives down construction costs, while online fuel reprocessing
reduces reactor downtime during reactor operation.

However, the liquid fuel form also brings about novel computational
challenges in simulating the transient behavior of \glspl{MSR}; the
interactions between neutronics and thermal-hydraulics are stronger due
to greater fuel material expansion. Furthermore, \glspl{DNP} in \glspl{MSR}
can flow freely within the primary coolant
loop as opposed to being held in place in a solid fuel matrix. Therefore,
the choice of coupling methods for each set of physics requires careful
consideration.

Most reactor analysis applications are usually reactor-specific by
design such as PARCS and TRACE for \glspl{LWR}, and SAS4A/SASSYS-1 for
liquid metal cooled reactors. Thus, these applications generally would not
account for \gls{MSR}-specific phenomena and are not suitable for \gls{MSR}
analysis without any modifications to the source code. Some research efforts
do focus on adapting these applications for \gls{MSR} analysis. Examples
include the coupling of modified versions of TRACE and PARCS
\cite{pettersen_coupled_2016}, and the development of VERA-MSR from the
integrated \gls{LWR} simulation tool VERA \cite{graham_development_2019}.
Others developed their \gls{MSR} simulation tools from general
multiphysics or \gls{CFD} applications such as COMSOL
\cite{fiorina_modelling_2014} and OpenFOAM \cite{aufiero_development_2014}.

This thesis demonstrates latest capabilities of Moltres
\cite{lindsay_introduction_2018}, an open source MSR simulation tool. Moltres
is an application built in the \gls{MOOSE} parallel finite element framework.
It solves
the deterministic multigroup neutron diffusion and thermal-hydraulics
\glspl{PDE} simultaneously on the same mesh. It supports up to 3-D meshes and
scales well for a large number of processors. The underlying \gls{MOOSE}
framework
provides a range of implicit and explicit methods for the coupling between the
neutronics and thermal-hydraulics governing equations.

In the introductory journal article for Moltres, Lindsay et al.
\cite{lindsay_introduction_2018} demonstrated Moltres' capabilities with
2D-axisymmetric and 3D models of the \gls{MSRE}. The results showed good
qualitative agreement with legacy \gls{MSRE} data but there were some minor
quantitative discrepancies due to a number of differences in the legacy model. 
Moltres has undergone further development in the past two years. The authors
from the first paper have since developed various new capabilities in Moltres
such as support for looping delayed neutron precursors back into the core,
and a heat exchanger object to set the heat removal rate.

In addition to those new capabilities, this thesis presents two more recent
developments in Moltres, namely fully integrating \gls{MOOSE}'s incompressible
Navier-Stokes module into Moltres, and introducing a
decay heat model. The incompressible Navier-Stokes module provides a more
accurate representation of the flow profile, precursor movement, and heat
transfer as opposed to assuming uniform or function-based velocity fields.

\section{Objectives}

The main objectives of this thesis are to demonstrate and verify Moltres'
latest capabilities in modeling multiphysics, steady-state, and transient
behavior of fast-spectrum \glspl{MSR} through the study of the \gls{MSFR}
concept. Code-to-code verification is an important exercise in software
development for ensuring that the application produces accurate and reliable
results. This thesis covers the \gls{MSFR} concept mainly because 
there is readily available literature data to verify our results against. The
steady-state and transient results also feature a significant degree of
physics interaction between the neutronics and thermal-hydraulics. This
present work will first present a verification of Moltres' \gls{MSFR}
neutronics results against Serpent 2, followed by a verification of
the coupled neutronics/thermal-hydraulics steady-state and transient accident
results against results from the Polimi and TUDelft models published by
Fiorina et al. \cite{fiorina_modelling_2014}. The
secondary objective is to identify areas of improvement in Moltres for future
research.

\section{Thesis Outline}

The outline of this thesis is as follows. Chapter 2 discusses the history and
features of \glspl{MSR}, and a literature review of existing \gls{MSR}
simulation tools. The chapter also covers the \gls{MSFR} concept in greater
detail. Chapter 3 details the simulation codes and the general modeling
approach for generating the results in this thesis. Chapter 4 provides a
neutronics assessment by comparing key neutronics parameters from Moltres'
eigenvalue calculations to Serpent's Monte Carlo calculations. Chapter 5
presents steady-state results of coupled neutronics/thermal-hydraulics
\gls{MSFR} simulations in Moltres. Chapter 6 presents transient accident
simulation results for unprotected reactivity insertions, unprotected loss of
heat sink, unprotected loss of flow, and unprotected pump overspeed,
respectively. Lastly, Chapter 7 summarizes the key findings in this thesis
and posits some potential avenues for future work.