### 2020-07-03 abstract.tex
more editing

This thesis presents the Molten Salt Fast Reactor simulation capabilities of
Moltres, a simulation tool for molten salt reactors.