### 2020-07-14 acks.tex
final comment edits

I wish to express my deepest gratitude to my advisor, Dr. Kathryn Huff, for
her guidance, patience, and encouragement in the past few years. My academic
journey thus far has been filled with many ups and downs and it would have
been much more difficult without her support. I would also like to thank Dr.
Tomasz Kozlowski for his help and invaluable comments as the second reader of
this thesis. I wish to acknowledge the funding support from the Singapore
Nuclear Safety \& Research Initiative since the start of my graduate studies.

I wish to thank my groupmates Dr. Andrei Rykhlevskii, Anshuman Chaube, and
Gwendolyn Chee for their input and ideas contributing to the completion of
this work. I also greatly appreciate Gavin Davis' and Matthew Kozak's help
with proofreading my writing.

Finally, no words can adequately express how thankful I am for my family's
unwavering support. I owe a lot to my parents and their efforts for all the
opportunities I've benefitted from.